Deriving Multi-headed Planar Dependency Parses from Link Grammar Parses

### 2014-05-24 doc/LinksToDAG.tex
updated the paper to include a table. Added some citations.

Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies. To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \TODO{}.

@@

We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph. This happened for roughly NNN\% of the corpus.}

@@

For example, setting the variables $$l_{LEFT} = TRUE, l_{RIGHT} = FALSE$$ would mean that the directional dependency of all links with label $l$ would be set to $LEFT$. If both variables were set to true, then the link type is allowed to go in either direction. Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.

@@

\subsection{Slack Hierarchy}

@@

\TODO{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information. This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.

This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.

\begin{equation}
  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
\end{equation}
\begin{equation}
  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
\end{equation}
Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
The objective is:
\begin{equation}
  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
\end{equation}
\begin{equation}
  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
\end{equation}

\begin{algorithm}
\caption{ILP Encoding}\label{ILP}
\begin{algorithmic}[1]
%\Procedure{MyProcedure}{}

\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
\State $\text{subject to}$
\State $\text{Acyclicity}$
\State $\text{All nodes reachable from the root.}$

%\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[ht!]
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{Precision and Recall of different runs on English}
\end{figure}

@@

\section{Link Corpus}

@@

@@

\appendix
\section{English Link Types}

>>>>

Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies.

To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. 
It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \TODO{}.

Kromann 2003 Danish dependency treebank\cite{kromann2003danish}

McDonald and Pereira 2006\cite{McDonald2006}

Sagae and Tsujii 2008\cite{Sagae2008}

Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}

@@

We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph. This happened for roughly 19\% of the sentences in the corpus.}

@@

For example, setting the variables $$l_{\textsc{LEFT}} = \textsc{TRUE}, l_{\textsc{RIGHT}} = \textsc{FALSE}$$ would mean that the directional dependency of all links with label $l$ would be set to $\textsc{LEFT}$. If both variables were set to true, then the link type is allowed to go in either direction. Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.

@@

\subsection{Slack}

@@

\begin{figure*}
%\centering

\begin{equation}
  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} link(a,L)
\end{equation}
\begin{equation}
  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
\end{equation}
%Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
OBJECTIVE:
\begin{equation}
  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
\end{equation}
\begin{equation}
  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
\end{equation}

\caption{The ILP formulation. L is a label and M is the size of the corpus.}

\end{figure*}

@@

\begin{figure}[ht!]
  \small
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{\small Precision and recall of increasing runs of English sentences}
\end{figure}

@@

\section{Link Corpus Experiment and Results}

\begin{figure}[ht!]
  \centering
  \small
  \input{figure/conll_analysis_sentences.tex}
  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
\end{figure}

@@

\section{Future Work}

\subsection{Slack Hierarchy}
We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information. This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types. We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.

@@

\clearpage
\appendix
\section{English Link Types}
